---
title: 'Probability and Statistics.'
authors: Hrebeniuk Dmytro, Baranovska Tetiana, Lakoma Nadiia
output:
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: 72
---

### Student\`s research "Human Stress Detection in and through Sleep by monitoring physiological data."

Analysing data from
<https://www.kaggle.com/datasets/laavanya/human-stress-detection-in-and-through-sleep?select=SaYoPillow.csv>
and predicting how different indicators affect the quality of our sleep.

#### Data overview:

Dataset includes information about every participant and their sleep
indicators: snoring rate, sleeping hour (\>= 2), heart rate, stress
level. Our aim is to find how these indicators affect each other and
whether there is a certain correlation or dependence between them.

sr - snoring rate

sh - hours of sleep

hr - heart rate

sl - stress level (0-4)

```{r}
pillow <- read.csv(file = 'C:/Users/baran/Desktop/data_for_research.csv')
pillow
```

```{r}

plotting <- function(cur_dat, bin, label){
  xrange = c(min(cur_dat),max(cur_dat))
  hist(cur_dat,
       probability = T, 
       breaks = bin, 
       xlim = xrange,
       main = label,
       col = "lightblue");
  lines(density(cur_dat),xlim=xrange, col = "blue")
  x <- seq(min(cur_dat),max(cur_dat), by=.01)
}
```

```{r}
plotting(pillow$sr, 20, "Snorring rate")
plotting(pillow$sh, 20, "Hours of sleep")
plotting(pillow$hr, 20, "Heart rate")

```

As we can see, all outcomes are equally possible.

**Snoring rate and heart rate.**

We want to find out how snoring during sleep affect heart rate.

```{r}
cor.test(pillow$sr, pillow$hr, method = "pearson")
```

A correlation coefficient tells the strength and direction of a
relationship between variables. Correlation coefficient is 0.9753004. We
have positive correlation coefficient close to 1, which tells us that
the higher snoring rate - the higher the heart rate and otherwise. That
means that when one variable changes, the other variables change in the
same direction.

We made a linear model in order to find the linear dependence between
snoring rate and heart rate and confirm the results with correlation
test.

```{r}
linear_model <- lm(pillow$sr ~ pillow$hr, data = pillow)
plot( pillow$hr, pillow$sr, pch=1, xlim=c(50,100), col = "blue")
abline(linear_model, col = "purple")
summary(linear_model)
```

Indeed, linear model gave the same result that the correlation test.
That is why we can say that there is positive dependence between snoring
rate and heart rate.

**Sleeping hours and stress level.**\
We cut all data where sleeping hours were less then 1 hour of seeping,
because we assume that this data is mistakenly taken.

```{r}
pillow$more <- pillow$sh > 5.0
```

Second hypothesis we test is\
$$
H_0: mean(\textrm{stress level of people sleeping less then 5 hours})  = \\=mean(\textrm{stress level of people sleeping more then 5 hours})
$$ $$
H_1:  mean(\textrm{stress level of people sleeping less then 5 hours})  >\\> mean(\textrm{stress level of people sleeping more then 5 hours})
$$

We used $t-test$ because the variance is unknown

1)  Extract our data

```{r}
more_sleeping <- c()
less_sleeping <- c()

for (i in 1:length(pillow$sh)){

  if (pillow$sh[i] > 5.0){
      more_sleeping <- append(more_sleeping, pillow$sl[i])
  }
  else {
     less_sleeping <- append(less_sleeping, pillow$sl[i])
  }
}
```

2)  hist to enforce our assumption

```{r}
hist(more_sleeping, breaks = 4)
hist(less_sleeping, breaks = 4)
```

We enforced our assumption that people who sleep longer have lower level
of stress

3)  $T-test$

```{r}
t.test(more_sleeping, less_sleeping, alternative = "greater")
```

We will not reject our $H_0$ hypothesis because our $p-value$ is
$\approx0.73$, which is very high and the higher p-value - the bigger
mistake when we reject $H_0$ .

Then we can say the test confirmed that actually such factor as sleeping
hours does not affect level of person's stress as average stress level
of those who sleep less than 5 hours equal to average level of stress of
those who sleep more than 5 hours.
