---
title: 'Probability and Statistics.'
authors: Hrebeniuk Dmytro ,Baranovska Tetiana, Lakoma Nadiia
output:
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: 72
---

### Student\`s research "Human Stress Detection in and through Sleep by monitoring physiological data."

Analysing data from
<https://www.kaggle.com/datasets/laavanya/human-stress-detection-in-and-through-sleep?select=SaYoPillow.csv>
and predicting how different indicators affect the quality of our sleep.

#### Data overview:

Dataset includes information about every participant and their sleep
indicators: snoring rate, sleeping hour (\>= 2), heart rate, stress
level. Our aim is to find how these indicators affect each other and
whether there is a certain correlation or dependence between them.

sr - snoring rate

sh - hours of sleep

hr - heart rate

sl - stress level (0-4)

```{r}
pillow <- read.csv(file = 'C:/Users/dimit/Downloads/data_for_research.csv')
pillow
```

```{r}

plotting <- function(cur_dat, bin, label){
  xrange = c(min(cur_dat),max(cur_dat))
  hist(cur_dat,
       probability = T, 
       breaks = bin, 
       xlim = xrange,
       main = label,
       col = "lightblue");
  lines(density(cur_dat),xlim=xrange, col = "blue")
  x <- seq(min(cur_dat),max(cur_dat), by=.01)
}

```

```{r}
plotting(pillow$sr, 20, "Snorring rate")
plotting(pillow$sh, 20, "Hours of sleep")
plotting(pillow$hr, 20, "Heart rate")

```

As we can see, all outcomes are equally possible.

```{r}
linear_model <- lm(pillow$sr ~ pillow$hr)
plot(pillow$sr, pillow$hr, pch=1, xlim=c(50,100), col = "blue")
abline(linear_model, col = "purple")
summary(linear_model)
```

Sleeping hours and stress level.\
We cut all data where sleeping hours were less then 1 hour of seeping,
becouse we assume that this data is mistakenly taken.

```{r}
pillow$more <- pillow$sh > 5.0

```

Second hypothesis we test is\
$$
H_0: mean(\textrm{stress level of people sleeping less then 5 hours})  = \\=mean(\textrm{stress level of people sleeping more then 5 hours})
$$ $$
H_1:  mean(\textrm{stress level of people sleeping less then 5 hours})  >\\> mean(\textrm{stress level of people sleeping more then 5 hours})
$$

We used $t-test$ because the variance is unknown

1)  Extract our data

```{r}
more_sleeping <- c()
less_sleeping <- c()

for (i in 1:length(pillow$sh)){

  if (pillow$sh[i] > 5.0){
      more_sleeping <- append(more_sleeping, pillow$sl[i])
  }
  else {
     less_sleeping <- append(less_sleeping, pillow$sl[i])
  }
}
```

2)  hist to enforce our assumption

```{r}
hist(more_sleeping, breaks = 4)
hist(less_sleeping, breaks = 4)
```

We don't enforced our assumption that people who sleep longer have lower
level of stress. But to be more sure we should perform t-test

3)  $t-test$

```{r}
#?t.test
t.test(more_sleeping, less_sleeping, alternative = "greater")
```

We will not reject our $H_0$ hypothesis because our $p-value$ is
$\approx0.73$
